\section{Limitations and Conclusion}
\label{sec:conclusion}

\paragraph{Limitations.} The binding limitation is that \emph{we have no human validation}. Our
``gold'' is a single, unvalidated LLM judge, so the alternative-annotator test
\citep{calderon2025alttest} is not actually run and the PPI/DSL guarantees
\citep{angelopoulos2023ppi,egami2023dsl} do not formally apply; every support-rate number is conditional
on that judge. Compounding this, the judge and the upstream claim/reference steps are all LLMs, so
\emph{self-preference} \citep{panickssery2024llm} may inflate the support rate and we cannot rule it out
without a non-self anchor. Beyond these: \emph{(3) stance is a proxy}---``support'' read from a title and
a ${\le}600$-character abstract is a coarse stand-in for a methods-aware reading; \emph{(4) retrieval
bounds the ceiling}---we score only relevance-retrieved evidence with no disconfirming search, so the
support rate is upward-biased and the refute rate a lower bound, and some ``references'' are bill text or
government pages, not science; \emph{(5) the gold set is small} ($n=120$) and from one model, so the
intervals are wide \emph{and} understate uncertainty about the gold itself; \emph{(6) strategic
text}---press releases are advocacy, so ``support'' for a chosen claim is not balanced evidence use;
\emph{(7) scope}---one chamber of one heavily-Democratic state, in English.

\paragraph{What a valid version would require.} The path from this feasibility scaffold to a valid
measurement is concrete: (i) a \emph{human}-coded anchor of roughly 100--200 claim--reference pairs,
sampled by a known design, against which to run the alt-test and to seed DSL/PPI; (ii) a
properly-configured claim-verification baseline (fine-tuned, rationale-based, in the SciFact line
\citep{wadden2020fact}) rather than zero-shot entailment, so that ``NLI'' is given a fair test; (iii) a
\emph{non-self} judge---a different model family, or human coders---to break the all-LLM circularity and
quantify self-preference; (iv) full-text, methods-aware stance rather than abstract-only; and (v)
disconfirming retrieval. None of these is exotic; together they would convert the conditional estimates
here into validated ones, and the released scaffold is built so that adding a human anchor is a drop-in
change.

\paragraph{Conclusion.} We asked whether claim--evidence \emph{support} can be measured at scale with no
human coders. Our answer is a qualified ``not yet, but here is a usable scaffold.'' We can build the
autonomous pipeline---claim extraction, retrieval, and an LLM judge anchoring a cheap NLI surrogate via
PPI---and run it end-to-end on open models with no API key; doing so yields a genuine cautionary result
(off-the-shelf zero-shot NLI is badly miscalibrated for this task) and a robust descriptive finding
(asymmetric AI \emph{engagement} across parties). But the headline support rates are conditional on an
unvalidated, possibly self-preferring judge, and we decline to present them as established. The
contribution is a reproducible template and an honest map of what stands between it and a valid
measurement, released so that others can supply the missing human anchor and put it to the test.

\paragraph{Reproducibility and ethics.} All code, the claim--reference data, the LLM-judge gold
labels, and the figure scripts are available at
\url{https://github.com/bdesmarais/ai-policy-science-use}; every number in this paper is reproducible
with a single command and no paid API. The corpus consists of public communications by elected
officials acting in their official capacity; we make no claims about private individuals. Because
automated ``evidence scoring'' of political speech could be misused to lend false objectivity to
partisan judgments, we stress that our estimates are measurements with stated uncertainty and known
biases, not verdicts, and that the cross-method diagnostics and human-anchor option are integral to
using the tool responsibly.
