\section{Discussion}
\label{sec:discussion}

We set out to measure, at scale, whether the scientific record corroborates the empirical claims
legislators make, and to do so in a way that can be \emph{trusted} without new in-domain human labels.
The result is partial, and we state it as such. The hardest \emph{judging} step can be validated by
transfer from established benchmarks: the high-capacity judge agrees with the expert gold at a level
approaching the human inter-annotator ceiling on SciFact ($\kappa=0.71$ vs the benchmark's $0.75$) and,
more tellingly, holds up on contested Climate-FEVER claims where the cheap surrogate drops to the
lexical floor. Because the benchmark text is human-written, this is not self-preference, and the judge holds up not only
on curated pairs but on \emph{retrieved}-evidence pairs that share the application's structure (AVeriTeC),
within the band of human disagreement. What keeps the \emph{pipeline} short of a fully closed instrument,
and forms the spine of the honest account below, is narrower than it once was: a single frontier judge
still leads on the hardest pairs (though \emph{combining} the open models narrows even that gap to within
sampling noise), and although every stage is independently validated against human labels, the stages'
\emph{joint} behaviour on a fresh in-domain sample is not separately audited.

\paragraph{Self-preference: bounded on the evidence side, open on the claim side.} The most serious
objection to an all-LLM loop is that a model judging model-surfaced evidence may exhibit self-preference
rather than assess science \citep{panickssery2024llm}. Benchmark validation defuses the
\emph{evidence}-side channel directly: SciFact and Climate-FEVER evidence is human-written, so a judge
that labels it correctly is reading the science, not recognizing its own prose. The cross-developer
panel \emph{corroborates} this---small open models from three developers also do the task above
chance---but we do not lean on it as an independent guarantee, because models trained on overlapping
corpora and aligned by similar methods do not err independently, and their agreement with the gold is
itself only modest ($\kappa=0.28$--$0.44$). One channel remains genuinely open: in the application the
\emph{claims} are LLM-extracted text, so the judge could favour model-generated phrasing on the claim
side even when the evidence is human-written, and the human-claim benchmarks do not test this.
Self-preference is therefore bounded on the evidence side but not closed on the claim side, and we say
so rather than implying the benchmarks settle it.

\paragraph{The pair-structure gap is substantially narrowed---against existing human labels.} The
standing objection was that our anchors paired a claim with evidence \emph{curated} to be about it
(SciFact, cited abstracts), whereas the application reads claims against \emph{retrieved} evidence that may
be off-topic; benchmark accuracy, the worry went, might not transfer across that gap in \emph{pair
structure}. Section~\ref{sec:res-application-validation} answers this on the pair structure the application
produces. AVeriTeC \citep{schlichtkrull2023averitec} pairs real-world claims with evidence \emph{retrieved
from the open web}---the application's pair structure---and the judge agrees with its trained-human
verdicts at $\kappa=0.75$ against the annotators' own pairwise $\kappa=0.62$; Climate-FEVER, whose evidence
is likewise retrieved rather than cited, tells the same story ($\kappa=0.67$ vs human $\alpha=0.33$). We
are careful not to over-read these: the judge's $\kappa$ is against the adjudicated gold and the human
figure is pairwise, so the honest reading is that the judge sits \emph{within the band of human
disagreement} by the alternative-annotator standard \citep{calderon2025alttest}, not that it ``exceeds''
the humans. Two further caveats keep this to ``narrowed'' rather than ``closed'': AVeriTeC is general
fact-checking rather than legislative AI, and its evidence was assembled to be adjudicable, whereas the
application's keyword hits are frequently off-topic (the retrieval misses documented in
\S\ref{sec:res-application} are exactly that). What is established is that the judge is within the human band
on retrieved-evidence pairs---the case that is hardest to validate
(Fig.~\ref{fig:band})---obtained, like the rest of our validation, from existing human labels with no
in-domain coding.

\paragraph{Why we do not gate on a fresh human audit.} One can still ask for a
few-hundred-pair human audit of \emph{our} corpus. We have not made it the gate, not from expedience but on
principle: a small human audit is not the clean ground truth the request assumes, for three reasons.
\emph{First, human coders are a noisy instrument on exactly these claims.} The benchmarks make this
measurable: on contested, lay claims of the kind legislators make, trained annotators agree only
weakly---Climate-FEVER at $\alpha=0.33$, AVeriTeC at $\kappa=0.62$. An audit of our pairs would inherit
that ceiling; the ``gold'' it produced would disagree with itself on a large share of contested cases, and
a judge--human disagreement there could equally be coded human--human. Measuring a low-noise instrument
against a high-noise one does not adjudicate the low-noise one. The right question is not ``does the judge
match a human gold'' but ``is the judge inside the band of human disagreement,'' which is the
alternative-annotator standard the field uses to license replacing a coder \citep{calderon2025alttest},
and which the judge already passes on application-structure pairs. \emph{Second, on partisan text an audit
reintroduces the bias it is meant to catch.} The sharpest worry about an LLM judge of political claims is
documented party-cue sensitivity---models shift labels when a party is named \citep{clemm2025partycues}.
Our design blunts this at the root: the judge never sees party. Its input is the claim and the reference's
title and abstract; ``Republican'' and ``Democrat'' appear nowhere in its prompt, so an explicit cue
cannot move it, and any residual partisan signal must travel through the claim's own wording, which a human
coder reads too. Human coders bring their own priors to partisan material, and political-science work now
finds LLMs matching or exceeding trained and outsourced human coders on precisely such tasks
\citep{tornberg2024llmexpert,heseltine2024llm}; a small RA-coded audit of partisan claims is not a
bias-free benchmark but another rater with a different, possibly larger, bias. We rest the case for not
gating on the audit on these two arguments and on the AVeriTeC evidence, and we deliberately do \emph{not}
lean on a further argument---that model labels can be regenerated whereas a human
gold cannot. A validation set is frozen by nature, and our case already rests on three frozen human-labeled
benchmarks whose labels no one can regenerate either, so regenerability does not distinguish benchmark
transfer from a frozen in-domain audit, and we drop it as a reason. What the no-human design does buy is
not a \emph{validation} advantage but a \emph{deployment} one: the instrument keeps pace with the
world---we refreshed twelve months of releases and recomputed the estimate in a sitting, which human coding
cannot match---and every label, query, and prompt is committed and auditable.

\paragraph{What an audit would still add, honestly.} None of this makes a human audit worthless. A large,
well-designed, \emph{multi}-coder audit---not a few hundred pairs---would add an independent in-domain
check and could tighten the bound, and we would welcome one. The one place a small human check was still
clearly worth having was \emph{retrieval relevance}---the step a stance audit does not cover, and the one
judgment on which humans are reliable---and we supply it by transfer:
the model's relevance judgment matches SciFact's human relevance labels at $96\%$ with perfect precision
(\S\ref{sec:res-retrieval}). That check certifies the easy discrimination (cited versus unrelated) and
rules out the worst failure mode; the harder retrieved-but-adjacent boundary, where the application's
misses live, is the one the judge is least certain on, shown by its concentration of errors in AVeriTeC's
\textsc{not-enough-evidence} class, and a cited-versus-same-topic test is the clean way to close it. What
remains for a human audit is
the \emph{joint} behaviour of extraction, retrieval, and judging on this specific corpus; its marginal
value, given that each stage is now checked against human labels and the judge sits within the
human-agreement band on application-structure pairs, is smaller than it might appear, it is
itself bounded by low human agreement on contested claims, and it would not make the deployed pipeline
autonomous, which is the contribution. We present the band evidence (Fig.~\ref{fig:band}) and the
stage-wise human-anchored checks as the validation standard appropriate to an instrument built to run
without people, and a fresh joint audit as a complementary check we did not make the gate---strong
evidence, we are clear, but not a complete substitute for one.

\paragraph{A reusable methodology, honestly scoped.} What generalizes beyond AI policy is the
\emph{methodology}: validate the judging step cheaply by benchmark transfer, propagate a small set of
expensive frontier-judge labels through a free surrogate by prediction-powered inference, and ground
references in an open index so the evidence base is auditable rather than fabricated. The cost profile
this yields is deliberate and, we think, defensible: a \emph{single} frontier model performs the language
steps---it extracts claims, guides retrieval over the free OpenAlex index (\S\ref{sec:res-retrieval}), and
judges stance---while the surrogate, the open combiners, the index, and the estimator cost nothing. References are real
records selected by the model from a free index rather than text it generated, so there is no
citation-fabrication exposure. Two honest qualifications travel with the methodology. A single open model
is competitive with the frontier judge only on \emph{clean} pairs, but \emph{combining} the open models
(a stacker trained on public benchmark labels, no in-domain labels) reaches the judge on SciFact and
narrows the Climate-FEVER gap to within sampling noise, so even the one paid component is, on the
judging step, largely replaceable. And while each stage is now independently validated against human
labels---extraction against check-worthiness labels, stance against curated \emph{and}
retrieved-evidence benchmarks, retrieval relevance against human relevance labels---what remains
unaudited is the stages' \emph{joint} behaviour on a fresh in-domain sample, which we argue is a
complementary check rather than the decisive gate.

\paragraph{Diagnose across methods.} A practical lesson concerns NLI as a stand-alone measure. On
clean benchmark pairs our zero-shot NLI classifiers are strong, but on the policy corpus they label
most relevance-retrieved references ``silent,'' because an abstract rarely \emph{entails} a
press-release claim in the strict textual sense even when it substantiates it. The discrepancy is not
a defect of the judge but a property of zero-shot entailment under relevance retrieval, and it is why
we use the validated judge as the anchor and NLI as a surrogate corrected by prediction-powered
inference rather than as a free-standing measure. Reporting cross-method agreement remains a useful
diagnostic, but it is the comparison to a validated standard that resolves which method to trust.

\paragraph{Limitations.} Beyond the frontier-judge and pair-structure issues above, several limits
temper the substantive findings. (1) \emph{Retrieval precision shapes the level, and is handled within
the single-model design.} The measured corroboration level is highly sensitive to retrieval
precision---broad keyword retrieval surfaces little claim-specific evidence and drives nearly all pairs to
\textsc{silent} (\S\ref{sec:res-robustness}). The model-guided retrieval of \S\ref{sec:res-retrieval}
addresses this, lifting relevance yield from $0.40$ to $0.97$ over the free index, reproducibly. Two
qualifications remain: stance is assessed only over \emph{relevant} retrieved references with no
deliberately disconfirming search, so corroboration is an upper and refute a lower bound; and the retriever
seeks relevant rather than disconfirming evidence, a directional caveat the validation cannot remove. (2)
\emph{Claim extraction is validated, not assumed.} Section~\ref{sec:res-extraction} establishes this with
no human coders. The extractor
matches human check-worthiness labels at the CheckThat! state of the art ($F_1$ $0.74$; $\kappa=0.53$ is
moderate, not human parity), and $95\%$ of its claims are entailed by their source release, so fabrication
is rare. Yield needs care because conditional and unconditional rates differ: \emph{given} a release with
empirical content the extractor is party-neutral ($7.1$ vs $7.3$ claims per release), but
\emph{unconditionally} Democratic releases yield more ($2.9$ vs $0.9$) because Republican AI releases more
often carry no empirical claim to extract. So extraction itself is not biased, but the Republican
corroboration arm is thin---$11$ empirical claims with retrievable references---so although its rate
($0.64$) is statistically indistinguishable from the Democratic rate ($0.60$), we read the cross-party
\emph{comparison} as suggestive rather than tested and rest the partisan finding on engagement volume,
which is robust. The residual extraction limitation is recall, not bias: at precision $0.83$
the extractor is deliberately conservative, so the corpus under-counts borderline empirical claims rather
than fabricating them. (3) \emph{Refute capability is real but nearly idle in the
application.} The judge's strong \textsc{refute} detection on the benchmarks answers the worry that the
pipeline cannot find disconfirming evidence, but relevance-seeking retrieval suppresses such evidence by
construction; it is the retrieval design, not the judge, that limits refute rates here. (4)
\emph{Engagement asymmetries lack member-level uncertainty.} The chamber is lopsidedly Democratic and a
few prolific offices can drive per-document intensity; the differences in Figure~\ref{fig:engage} are
point estimates over unweighted aggregates and should be read as descriptive, with member-clustered
intervals a needed refinement. (5) \emph{Strategic text and scope}: press releases are advocacy, so a
high corroboration rate reflects selective citation as much as evidence quality; and the application is
one chamber of one state, with claim extraction still unvalidated even though the judging and retrieval
steps are now characterized, so multi-state deployment inherits that gap rather than only requiring more
corpus.

\paragraph{Ethics.} Automated ``evidence scoring'' of political speech could lend false objectivity to
partisan judgments. We stress that our outputs are measurements with stated uncertainty and known
bounds, not verdicts; that they concern public communications by officials acting in their official
capacity; and that the validation, cross-method diagnostics, and open code are integral to using the
instrument responsibly.
