\section{Introduction}
\label{sec:intro}

Adjudicating whether a body of evidence \emph{supports}, \emph{refutes}, or is \emph{silent} on a specific
empirical claim is one of the most demanding measurement tasks in the quantitative social and information
sciences. It is the judgment behind automated fact-checking, systematic review and evidence synthesis,
misinformation detection, and the study of how science enters public decisions---and in each it has
historically required trained human coders reading both the claim and the underlying science
\citep{grimmer2013text}. Expert coding is slow, expensive, and hard to reproduce, so work of this kind has
been confined to small samples or coarse proxies and cannot keep pace with the volume of modern
communication. A general, \emph{trustworthy}, and reproducible way to make this judgment at scale would
matter wherever text must be checked against evidence.

We build such an instrument and demonstrate it on a question of independent importance: whether the
scientific record corroborates the empirical claims that legislators make, a central concern for the study
of evidence-based policymaking \citep{oliver2022works,bogenschneider2019revisiting,weiss1979many}. The
choice is deliberate---political claims are contested, the evidence is retrieved rather than cited, and the
stakes of a partisan measurement error are high---so a method that earns trust here transfers readily to
the easier settings above.

Large language models (LLMs) appear to dissolve this bottleneck. Across political science and
computational social science, instruction-tuned LLMs now match or exceed trained human coders on stance
detection, framing, and other annotation tasks, at a fraction of the cost
\citep{gilardi2023chatgpt,ziems2024can,heseltine2024llm,tornberg2025llm}. This raises the possibility of
an \emph{autonomous} measurement pipeline---one that ingests raw political text, extracts the empirical
claims, retrieves topically-relevant science, and judges whether that science supports each claim. The
aspiration is to do this with no human annotators and without a proliferating stack of paid services. We
take a deliberately pragmatic position on cost: one frontier model performs the language steps, and
everything around it---the bibliographic index, the scalable surrogate, the estimator---is free and open.
The constraint we hold to is not ``no proprietary model'' but ``no \emph{second} paid service,'' and we
ask how much of the pipeline can meet that bar. The answer turns out to be most of it.

The obstacle is not building such a pipeline but \emph{trusting} it. If the model that judges
claim--evidence support is never checked against a human standard, its outputs are unfalsifiable: a
fluent label is not necessarily a correct one \citep{bender2021stochastic}, and the statistical
machinery for turning noisy machine labels into valid population estimates---design-based supervised
learning and prediction-powered inference \citep{egami2023dsl,angelopoulos2023ppi}, and the
alternative-annotator test that licenses replacing a human coder \citep{calderon2025alttest}---is all
defined relative to trusted human ground truth. A second, subtler threat is circularity: when the
claims, the retrieved references, and the judge are \emph{all} produced by language models, an LLM
judge may rate model-surfaced evidence as supportive because LLM evaluators are known to prefer
model-generated and low-perplexity text---a documented \emph{self-preference} bias
\citep{panickssery2024llm,zheng2023judging}. An autonomous pipeline that is merely reproducible
reproduces these biases exactly.

Here we show that this trust gap can be closed without collecting new human annotations in the target
domain, by \emph{validating the autonomous judge against existing human-labeled benchmarks for the
identical task}. Scientific claim verification---deciding whether an abstract supports, refutes, or is
neutral toward a claim---is precisely the support-assessment step of our pipeline, and it has mature,
expert-annotated benchmarks: SciFact \citep{wadden2020fact} for expert scientific claims and
Climate-FEVER \citep{diggelmann2020climate} for contested, real-world claims phrased in lay language.
A judge that matches the expert gold on these benchmarks---and approaches the human inter-annotator
agreement the benchmarks themselves report---has demonstrably learned the task; and because the
benchmark text is human-written rather than model-generated, accuracy there cannot be an artifact of
self-preference on the evidence side. Benchmark validation thus supplies both the human anchor that
downstream inference requires and a direct test of the circularity threat. The two benchmarks bracket
our target: SciFact is cleaner and more technical than legislative claims, Climate-FEVER is noisier,
contested, and lay---so performance across both characterizes the judge where policy communication
falls.

Using this strategy we study a measurement pipeline---claim extraction, reference retrieval from the
open OpenAlex index \citep{priem2022openalex}, and stance judging---applied to a decade of legislative
communication on artificial intelligence (AI). We make six contributions, and we scope each carefully.
(1) A transferable \emph{validation methodology}: benchmark-anchored evaluation that tests whether an
automated stance judge can be trusted with no new in-domain human labels, by scoring it on existing
human-labeled claim-verification benchmarks for the identical task. (2) A measurement finding about open
versus closed models. Single open models that run locally trail a proprietary frontier judge---only the
NLI surrogate matches it, on clean SciFact pairs alone. But \emph{combining} the open models, with no new
in-domain labels, closes much of the gap: a stacker trained only on the public benchmark gold reaches the
judge on SciFact and narrows the harder-task gap to within sampling noise. So the \emph{judging} step
does not, after all, require a proprietary model---an important correction to a ``you need a closed
model'' reading. (3) A \emph{retrieval} method that is reproducible, claim-targeted, and DOI-grounded over a
free index. The frontier model reads each claim and, when the claim has a scientific core, emits a targeted
query against the free OpenAlex (or Crossref) index, abstaining otherwise; the result surfaces a genuinely
relevant reference for $97\%$ of empirical claims against $40\%$ for naive keyword search, abstains on every
non-scientific claim a keyword search blindly answers, and returns real, dereferenceable records---and the
relevance judgment itself matches human relevance labels at $96\%$, so retrieval is human-anchored rather
than self-graded. (4) An application that recovers a
robust partisan asymmetry in AI engagement and estimates the rate at which legislators' claims are
\emph{corroborated by the topically-nearest literature}---which is precisely what the instrument
measures, and which we distinguish from observed evidence \emph{use}, since legislators do not cite the
papers we retrieve. (5) A validation of the \emph{extraction} step, again with no human coders: the
extractor matches human check-worthiness labels (ClaimBuster) at the published state of the art, $95\%$
of its claims are entailed by their source release (so it rarely fabricates), and its yield is
party-balanced---so the benchmark-transfer philosophy that anchors the judge also anchors the front end.
(6) An explicit account of what is and is not validated: extraction, retrieval, and judging are each now
checked. Crucially, we validate the judge not only on curated pairs but on \emph{retrieved}-evidence,
application-structure pairs (AVeriTeC \citep{schlichtkrull2023averitec}), where it agrees with human
labels \emph{within the band of human disagreement}, and we validate the retrieval-relevance judgment
against human relevance labels ($96\%$ accuracy)---so we mount a principled case that a fresh in-domain
human audit is a \emph{complementary} check on the joint pipeline rather than the decisive gate the
no-human-coder design forecloses (\S\ref{sec:discussion}). The result is a validated
extraction-retrieval-judging pipeline whose every stage is checked without new human coding, and an honest
map of what an audit would, and would not, still add.

The contribution we most want to foreground is methodological and travels well beyond this application.
The central difficulty in deploying an autonomous LLM measurement instrument is not building it but
\emph{trusting} it without the human gold standard that licensing such instruments normally requires. We
show that benchmark transfer---validating each stage against \emph{existing} human-labeled data for the
identical sub-task, and quantifying where the validation does and does not reach---can substitute for fresh
in-domain coding, and that a party-blind design neutralizes the specific failure mode (partisan-cue
sensitivity) that makes LLM judging of political text hazardous. The same recipe applies wherever claims
must be measured against evidence at a scale human coders cannot meet and where a gold standard is scarce,
costly, or itself noisy: evidence synthesis, fact-checking, misinformation surveillance, and the empirical
study of science in society. The policy application is the demonstration; the transferable, honestly
bounded validation methodology is the result.
