\section{Methods}
\label{sec:methods}

\subsection{The pipeline}
\label{sec:methods-pipeline}
The pipeline maps a corpus of political documents to party-level estimates of claim--literature
corroboration in three stages---claim extraction, evidence retrieval, and support assessment---with no
human annotation in the loop. The language steps are performed by a single frontier model; the
bibliographic index, the scalable surrogate, the open-model panel, and the estimator are free and open.

\paragraph{Claim extraction.} Each document is processed by an instruction-tuned LLM prompted to
extract up to ten empirical claims---propositions ``for which, if you were writing a scientific paper,
you would need a reference''---as structured records (document id, claim number, claim text), isolating
verifiable propositions from rhetorical or procedural text \citep{grimmer2013representational}.

\paragraph{Evidence retrieval.} Retrieval is \emph{model-guided} search over a free bibliographic index:
the same frontier model that judges stance reads each claim and either emits a concise, claim-targeted
query or \emph{abstains} when the claim is legal, procedural, political, or a value statement with no
scientific literature to retrieve. The query runs against \emph{OpenAlex} \citep{priem2022openalex}, a
free open index of over 250 million works restricted to research articles with abstracts, or, equivalently,
against \emph{Crossref}, the free DOI registry; the approach is index-agnostic, and every returned
reference is a real, dereferenceable record, eliminating the citation-fabrication failure mode of asking a
model to ``name references.'' The model's queries and abstention decisions are committed as static data
(\texttt{claude\_retrieval.py}, \texttt{crossref\_retrieval.py}), so the evidence base is reproducible. We
compare model-guided retrieval against a \emph{naive keyword} baseline---the same index queried with
stopword-stripped claim keywords, which is reproducible but untargeted, cannot abstain, and returns broad,
topically-adjacent papers.

\paragraph{Retrieval evaluation.} We evaluate retrieval on a balanced 50-claim sample (25 Democratic and
25 Republican claims, of which the model judged 30 to have a scientific empirical core and 20 not). For
each arm we retrieve the top eight hits and measure three things. \emph{Relevance yield}: the fraction of
empirical claims whose top three references contain at least one the judge accepts as genuinely about the
claim. \emph{Abstention}: on the 20 non-empirical claims, whether each arm returns references it should
not. \emph{Verifiability}: the share of returned references carrying a DOI. We separately validate the
relevance judgment itself against human labels (\S\ref{sec:res-retrieval}). All relevance and stance
labels are committed (\texttt{data/retrieval\_relevance\_labels.json}).

\paragraph{Support assessment.} Each (claim, reference) pair is assigned a stance---\textsc{support},
\textsc{refute}, or \textsc{silent} (irrelevant or insufficient information)---by the validated judges
described next. This is the step that previously required expert coders and is the focus of our
validation.

\subsection{Validators}
\label{sec:methods-validators}
We treat stance assignment as an annotation problem and use validators that fail in different ways, so
that their agreement is informative \citep{zheng2023judging}.
\begin{itemize}
  \item \textbf{LLM judge (proprietary).} A high-capacity instruction-tuned model (Claude Opus~4.8),
  queried in context without fine-tuning, reads the claim and the reference title and abstract and
  returns a stance label. We are explicit that this is a \emph{proprietary} frontier model: it does not
  run locally, it is not free, and we do not present it as an open component (see disclosure). Models of
  this class match or exceed trained human coders on stance tasks
  \citep{gilardi2023chatgpt,heseltine2024llm,tornberg2025llm}; whether the \emph{open} alternatives below
  can replace it is one of our questions.
  \item \textbf{Open-source LLM panel.} Three instruction-tuned models from three independent
  developers---Qwen2.5-3B-Instruct (Alibaba), Phi-3.5-mini-instruct (Microsoft), and
  OLMo-2-7B-Instruct (Allen Institute for AI)---run the identical stance prompt locally. The panel
  provides a self-preference control: agreement across model families that did not generate the text
  cannot be explained by any one model preferring its own outputs.
  \item \textbf{NLI classifiers.} Two open-source natural-language-inference models with different
  architectures and training data---\texttt{DeBERTa-v3-base-mnli-fever-anli} and
  \texttt{bart-large-mnli}---score each pair with the reference as premise and the claim as hypothesis,
  mapping \textsc{entailment}$\rightarrow$\textsc{support},
  \textsc{contradiction}$\rightarrow$\textsc{refute}, \textsc{neutral}$\rightarrow$\textsc{silent}
  \citep{yin2019benchmarking,laurer2024less}. NLI is the standard discriminative approach to scientific
  claim verification \citep{wadden2020fact,koneru2024llm} and supplies a cheap surrogate that scales to
  every pair.
  \item \textbf{Baselines.} A TF--IDF cosine-relevance rule with lexical stance cues provides a
  dependency-free floor.
\end{itemize}
All of these \emph{open} models run locally on a single consumer GPU (Apple M-series, Metal backend)---the
proprietary judge, by contrast, does not run locally and is reached through a paid product (see
disclosure); the heaviest local component
is a 7B-parameter model in half precision.

\paragraph{Open combiners.} Because the open models fail in complementary ways (NLI is strong on clean
pairs and weak on lay ones; the panel is the reverse), we also evaluate three label-frugal combiners over
their committed per-pair predictions: a zero-shot \emph{majority vote} (no labels); a transfer-weighted
vote (per-class-$F_1$ weights estimated on the \emph{other} benchmark); and a \emph{stacker} (logistic
regression, with a gradient-boosted variant) trained on the \emph{public} benchmark gold with five-fold
out-of-fold cross-validation, text features fit inside each fold to prevent leakage, and model-vote plus
light lexical features. The stacker uses existing public labels but \emph{no new in-domain labels}---the
standard the rest of the pipeline holds to (\texttt{open\_ensemble.py}).

\subsection{Human-anchored validation}
\label{sec:methods-validation}
The central methodological move is to validate every judge against expert-human labels for the
\emph{identical} task before trusting it on the unlabeled policy corpus. We use two benchmarks. SciFact
\citep{wadden2020fact} comprises expert-written claims paired with cited abstracts and human
\textsc{support}/\textsc{contradict}/\textsc{noinfo} labels; we form one (claim, abstract) pair per
cited document, yielding 340 labeled pairs. Climate-FEVER \citep{diggelmann2020climate} comprises
real-world, often contested climate claims with retrieved evidence sentences, each human-labeled
\textsc{supports}/\textsc{refutes}/\textsc{not-enough-info}; we evaluate on a class-balanced sample.
Both label schemes map onto our \textsc{support}/\textsc{refute}/\textsc{silent} classes.

We run the NLI classifiers, the open-source panel, and the TF--IDF baseline on every benchmark pair
programmatically. The proprietary judge we use deliberately sparsely (mirroring its gold-on-a-sample
role): we draw a class-stratified sample, present each (claim, reference) pair \emph{blind to the gold
label}, and record its stance---$n=66$ pairs on SciFact and $n=45$ on Climate-FEVER. To keep accuracies
\emph{comparable}, we additionally evaluate every other validator on the \emph{exact} pairs the judge
saw, and report Wilson $95\%$ intervals throughout, because the judge's sample is small. We report
accuracy, macro-$F_1$, per-class $F_1$, confusion, and pairwise $\kappa$. Two scope points are built into
the design. First, ``human-level'' requires a human baseline, so rather than assert it we compare the
judge's agreement with the gold to the human inter-annotator agreement the benchmarks themselves report
(Cohen's $\kappa=0.75$ for SciFact; Krippendorff's $\alpha=0.334$ for Climate-FEVER). Second, these
benchmark pairs are \emph{curated to be about each other}---SciFact pairs a claim with its cited
abstract, Climate-FEVER with relevance-retrieved evidence---whereas in the application the judge reads
claims against keyword hits that may be off-topic. Benchmark accuracy therefore validates the
\emph{judging} operation, not its transfer to the noisier application pairings, a gap we treat as the
foremost limitation (\S\ref{sec:discussion}).

\subsection{Validating the extraction step}
\label{sec:methods-extraction}
Stance judging is the step the literature flagged, but \emph{claim extraction}---an instruction-tuned
model reading a release and emitting its empirical claims---was the step we had left unchecked. We
validate it with no new human coders, by three complementary tests that mirror the judge's
benchmark-anchored validation (\S\ref{sec:res-extraction}).
\emph{(i) Check-worthiness against human labels.} The extractor's core decision---``is this an empirical
claim for which one would need a reference?''---is the fact-checking task of \emph{check-worthiness
detection}, which has mature human-labelled benchmarks. We score the same models on ClaimBuster
\citep{hassan2017claimbuster}, whose ${\sim}23$k US-debate sentences are labelled check-worthy-factual,
unimportant-factual, or non-factual by trained humans, on a class-balanced sample (binary check-worthy
vs not), alongside the open-source panel and a lexical baseline, and we read accuracy against the human
inter-annotator and CheckThat! state-of-the-art ceilings \citep{barron2020checkthat}.
\emph{(ii) Faithfulness (grounding).} The gravest extraction failure is fabricating a claim the document
does not support. For every extracted policy claim we test that it is \emph{entailed by its own source
release} using natural-language inference---the standard out-of-the-box faithfulness metric
\citep{honovich2022true,laurer2024less}---splitting the release into sentence windows and taking the
maximum entailment; a claim is \emph{grounded} if some part of its source entails it. We then have the
validated judge review the lowest-entailment claims to confirm they are genuine fabrications rather than
paraphrase the entailment model missed.
\emph{(iii) Yield neutrality.} Finally we test whether extraction \emph{yield} differs by party in a way
that could confound the per-claim comparison, by comparing claims extracted per release across parties.

\subsection{Valid estimation under imperfect labels}
\label{sec:methods-ppi}
Plugging machine labels directly into a downstream analysis biases estimates and invalidates confidence
intervals \citep{egami2023dsl}. We therefore treat the scalable NLI labels as a \emph{surrogate}
available for all $N$ pairs and the LLM-judge labels as \emph{gold} on a sample of $n \ll N$ pairs, and
combine them with prediction-powered inference \citep{angelopoulos2023ppi}. For the binary support
indicator, the PPI estimate of the population support rate is
\begin{equation}
  \hat{\theta}_{\text{PPI}} = \underbrace{\frac{1}{N}\sum_{i=1}^{N} f(x_i)}_{\text{surrogate mean (all pairs)}}
  + \underbrace{\frac{1}{n}\sum_{j=1}^{n}\bigl(y_j - f(x_j)\bigr)}_{\text{rectifier (gold sample)}},
  \label{eq:ppi}
\end{equation}
with $f(x_i)\in\{0,1\}$ the NLI ``support'' prediction, $y_j\in\{0,1\}$ the LLM-judge gold, variance
$\widehat{\mathrm{Var}}=\mathrm{Var}(f(x))/N + \mathrm{Var}(y-f(x))/n$, and a Wald 95\% interval
$\hat{\theta}_{\text{PPI}}\pm 1.96\,\widehat{\mathrm{Var}}^{1/2}$. Equation~\eqref{eq:ppi} is the
prediction-powered special case of the design-based supervised learning estimator \citep{egami2023dsl};
both reduce to the human-coded estimate as $n\to N$. Two cautions apply to our application and we state
them plainly. First, the gold labeler is the LLM judge, so the intervals are valid for the rate the
\emph{judge} would assign over all pairs; their validity for a \emph{human}-equivalent rate rests on the
in-domain transfer we cannot yet test (\S\ref{sec:discussion}). Second, the efficiency gain from the
surrogate is small here: on the policy corpus the NLI labels are nearly constant (almost all
\textsc{silent}), so the surrogate variance term nearly vanishes and the estimate reduces to the judge's
mean on the gold sample. For the application we judge a stratified gold sample of $56$ claims spanning the
full AI corpus---$45$ Democratic claims drawn at random and all $11$ Republican empirical claims for which
the index returned references---and report Wilson intervals on the judge's labels
(\texttt{historical\_corroboration.json}). PPI thus functions as valid inference for the judge's labels on
a gold sample, not as a large variance reduction from a strong surrogate, and we present the precision
accordingly.

\subsection{Corpus and reproducibility}
\label{sec:methods-corpus}
The political application uses a corpus of California State Assembly member press releases, partitioned by
the issuing member's party, of which the AI-related subset is analysed here; a parallel K--12
education-policy corpus is used to test transfer of the extraction-and-attention front end. The corpus is
kept current: in addition to the historical 2015--2025 material we \emph{refresh} it through the most
recent twelve months by re-scraping both caucus archives---the Assembly Democratic Caucus
(\texttt{asmdc.org}) and the Assembly Republican Caucus (\texttt{asmrc.org})---from their published
sitemaps, the same uniform method for both parties (\texttt{fetch\_refresh\_releases.py}); the latest
window analysed here runs to June 2026 (\S\ref{sec:res-refresh}). The asymmetry between the two archives
is itself informative and we report it plainly: the Democratic archive is comprehensive and date-stamped,
while the Republican caucus site indexes far fewer releases and essentially no AI-focused ones in the
window.

\paragraph{Index portability.} Model-guided retrieval is not tied to one bibliographic index: the same
claim-targeted queries run against \emph{OpenAlex} or \emph{Crossref}---the free DOI registry, every record
a registered DOI by construction (\texttt{crossref\_retrieval.py})---and we report results under both,
which establishes that the method depends on the model's targeting rather than on any one service. Code,
the benchmark harness, the committed judge labels and queries, the retrieved references, the scraped
release text, and the figure scripts are public at
\url{https://github.com/bdesmarais/ai-policy-science-use-public}, and the analysis regenerates from
committed inputs with only the single frontier model.
