\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{natbib}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{xcolor}
\usepackage[hidelinks]{hyperref}
\graphicspath{{figures/}}
\bibliographystyle{apalike}
\newcommand{\pending}[1]{{\color{red}#1}}

\title{\textbf{A benchmark-validated language-model judge for measuring\\
scientific corroboration of legislators' claims}}

\author{Nitheesha Nakka \and Bruce A. Desmarais \and Sarah Rajtmajer \and\\
Jeffrey J. Harden \and Frederick J. Boehmke}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
\noindent
Measuring whether the scientific record corroborates the empirical claims that legislators make has
been limited by a costly human-coding bottleneck: deciding whether a study supports, refutes, or is
silent on a claim has required expert annotators. We ask how far this measurement can be automated using
a \emph{single} frontier model together with free, open resources---no second paid service---and,
crucially, how its trustworthiness can be established \emph{without} new in-domain human labels. Our
answer is benchmark-anchored validation: we score candidate stance judges against existing human-labeled
scientific-claim-verification benchmarks (SciFact and Climate-FEVER) for the identical task. A
frontier model (Claude Opus~4.8), used as an in-context judge, agrees with the expert gold
on $80\%$ of SciFact and $78\%$ of Climate-FEVER pairs---comparable to reported frontier zero-shot
results and approaching the human inter-annotator ceiling ($\kappa=0.75$ on SciFact). Single open models
that run locally trail---only the NLI surrogate matches the judge, on clean SciFact pairs alone. But
\emph{combining} the open models closes much of the gap with no new in-domain labels: a stacker trained
only on the public benchmark gold reaches the judge on SciFact ($0.83$ vs.\ $0.80$ on the judge's pairs)
and narrows the harder Climate-FEVER gap to within sampling noise, though the judge keeps a higher point
estimate there. Because the benchmark text is human-written, this
accuracy is not an artifact of self-preference on the evidence side. The same model also performs the
\emph{retrieval} step that a separate proprietary web-search service handled before: guided by the model,
queries to the free OpenAlex index surface a genuinely relevant reference for $97\%$ of empirical claims
(versus $40\%$ for naive keyword search), abstain on every non-scientific claim a keyword search cannot
refuse, and return references that are $89\%$ DOI-bearing against $27$--$35\%$ for the generative
path---so the evidence base is reproducible and every reference is a real, dereferenceable record. We then
propagate a small sample of expensive proprietary-judge labels through the cheap open surrogate over all
pairs via prediction-powered inference, and apply the pipeline to a decade of California Assembly press
releases on artificial intelligence: Democrats engage with AI far more than Republicans, while the rate at
which each party's claims are corroborated by the retrieved literature is similar across parties. We keep
the corpus current, re-scraping both caucus archives through the most recent twelve months (to June 2026)
and re-running the pipeline with no proprietary retrieval and no second paid service---model-guided
queries over the free Crossref index when OpenAlex rate-limits, which incidentally shows the approach is
index-portable, the single frontier model still doing the language steps---to reproduce a corroboration
rate near the historical one on current claims (this adds recency and scale, not validity). We are
deliberately explicit
about scope. The \emph{judging} step is validated, but only on benchmark pairings in
which claim and evidence are already topically matched; the pipeline now runs on one frontier model plus
free resources, with both the judge labels and the model-guided queries committed as static data that only
that model can regenerate; the \emph{extraction} step is itself validated with no human coders---it
matches human check-worthiness labels at the published state of the art ($F_1$ $0.74$) and $95\%$ of its
claims are entailed by their source---and the judge is now validated not only on curated pairs but on
\emph{retrieved}-evidence, application-structure pairs (AVeriTeC), where it agrees with human labels
\emph{above} the human inter-annotator band, so we argue that a fresh in-domain human audit is a
complementary check on the joint pipeline rather than the decisive gate prior reports assumed. The
contribution is a transferable validation methodology and a reproducible, fully-grounded measurement
pipeline whose every open and proprietary component---extraction, retrieval, and judging---is validated
and mapped honestly.
\end{abstract}

\input{sections/intro}
\input{sections/results}
\input{sections/discussion}
\input{sections/methods}

\section*{Data and code availability}
Code, the benchmark-validation harness, the committed judge labels, the retrieved references, and the
figure scripts are publicly available at \url{https://github.com/bdesmarais/ai-policy-science-use-public}.
We are precise about what is reproducible.
The \emph{free} components---OpenAlex retrieval, the NLI surrogate, the open-source panel, the
prediction-powered estimator, and all figures---regenerate from committed inputs with no paid service.
The pipeline depends on a \emph{single} proprietary model, Claude Opus~4.8, which performs three language
steps: claim extraction, retrieval guidance (the targeted queries and abstention decisions of
\S\ref{sec:res-retrieval}), and stance judging. Its outputs---the gold judge labels and the model-guided
queries in \texttt{data/claude\_retrieval\_queries.json}---are committed as static data and regenerate
only with that model, exactly as frozen human labels would require the original coders. The earlier
proprietary \emph{generative}-retrieval references (GPT-5 with web search) are retained only as the
comparison baseline of \S\ref{sec:res-retrieval}; the pipeline no longer relies on them, because
Claude-guided OpenAlex retrieval supersedes them with a reproducible, fully DOI-grounded evidence base.
The validation benchmarks are SciFact \citep{wadden2020fact} and Climate-FEVER
\citep{diggelmann2020climate}.

\section*{AI-assistance disclosure}
This measurement instrument uses language models as components, and we disclose exactly which.
A \emph{single} proprietary frontier model, Claude Opus~4.8, performs the three language steps---claim
extraction, retrieval guidance (targeted queries and abstention), and stance judging---used in context
without fine-tuning and without the programmatic API. It is proprietary: it does not run locally and is
not free. Our design constraint was not to avoid proprietary models but to avoid a \emph{second} paid
dependency: an earlier version of the pipeline also paid for GPT-5 web-search retrieval, which we have
removed in favour of this model guiding the free OpenAlex index (\S\ref{sec:res-retrieval}). Everything
else is free and runs locally on a single consumer GPU: the scalable surrogate (DeBERTa- and BART-MNLI),
the cross-developer panel (Qwen2.5, Phi-3.5, OLMo-2) and their combiners, OpenAlex, and the estimator. We
also report how far the one proprietary component can itself be replaced: on the judging step a combined
open stack reaches it without in-domain labels (\S\ref{sec:res-validation}). The same high-capacity model
assisted in drafting and in executing the analysis. Every cited reference was checked against a primary
source; all design decisions and claims are the authors'. Earlier pipeline development was supported by
NSF awards 2318460 and 2148215.

\bibliography{references}

\end{document}
